Generating Images From Captions\\ With Attention

### 2015-11-01 iclr-paper-new.tex
a

\section{Experiments}
\vspace{-0.05in}
\subsection{Microsoft COCO}
\vspace{-0.05in}
Microsoft COCO \citep{mscoco} is a very large dataset of roughly 83k images, each annotated with 5 captions. The rich collection of images with a wide variety of styles, backgrounds and objects makes the task of learning a good generative model 
%conditioned on a caption 
very challenging. For consistency with related work on caption generation, we used only the first five captions when training and evaluating our model. 
The images were resized to $32 \times 32$ dimension for consistency with other tiny image datasets \citep{krizhevsky_cifar}. In the following subsections, we analyzed both the qualitative and quantitative aspects of our model as well as compared its performance with that of other, related generative models\footnote{To see more generated images, go to \url{http://www.cs.toronto.edu/~emansim/cap2im.html}}.

@@

We compared the performance of the proposed model to the DRAW model conditioned on captions without $align$ function (noalignDRAW) as well as the DRAW model conditioned on skipthought vectors \citep{kiros_skipthought} (skipthoughtDRAW). All of the conditional DRAW models were trained with binary cross-entropy cost function. We also compared our model with Fully-Connected (Fully-Conn) and Convolutional-Deconvolutional (Conv-Deconv) Variational Autoencoders which were trained with the least squares cost function. The LAPGAN model of~\cite{denton_lapgan} was trained on a two level Laplacian Pyramid with a GAN as a top layer generator and all stages were conditioned on the same skipthought vector.

\begin{table}[!t]
\begin{center}
\begin{tabulary}{\linewidth}{c || c c c c c || c c}
\hline
\multicolumn{7}{c}{\textbf{Microsoft COCO (before sharpening)}} \\
\hline
& \multicolumn{5}{c||}{Image Search} & \multicolumn{1}{c}{Image Similarity} \\
%\cline{2-6}
\textbf{Model} & \textbf{R@1} & \textbf{R@5} & \textbf{R@10} & \textbf{R@50} & \textbf{Med r} & \textbf{SSI}\\ %& \textbf{Feature Distance} \\
\hline
\hline
LAPGAN & - & - & - & - & - & 0.08 \\ % & 6.89 \\ % - & - & - & - & - & 0.08
\hline
Fully-Conn VAE & 1.0 & 6.6 & 12.0 & 53.4 & 47 & 0.156 \\ %& 6.11 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
Conv-Deconv VAE & 1.0 & 6.5 & 12.0 & 52.9 & 48 & 0.164 \\ %& 6.32 \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
skipthoughtDRAW & 2.0 & 11.2 & 18.9 & 63.3 & 36 & 0.157 \\ %& 6.93 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
noalignDRAW & 2.8 & 14.1 & 23.1 & 68.0 & 31 & 0.155 \\ %& 7.01 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
alignDRAW & 3.0 & 14.0 & 22.9 & 68.5 & 31 & 0.156 \\ %& 6.99 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
\end{tabulary}
\end{center}
\caption{Retrieval results of different models. 
\textbf{R@K} is Recall@K
    (high is good). \textbf{Med} {\it r} is the median rank (low is
    good). 
}
\label{tab:results}
\vspace{-0.1in}
\end{table}

@@

As for the quantitative comparison of different models, we first compare the performances of the model trained with variational methods. We rank the images in the test set conditioned on the captions based on the variational lower bound of the log-probabilities 
and then report the Precision-Recall metric as an evaluation of the quality of the generative model (see Table~\ref{tab:results}.). Unsurprisingly, generative models did not perform well on the image retrieval task. To deal with the large computational complexity involved in looping through each test image, we create a shortlist of one hundred images including the correct one, based on the images having the closest distance in the convolutional feature-space of a VGG-like model \citep{simonyan_convnet} trained on the CIFAR dataset\footnote{The architecture of the model is described here \url{http://torch.ch/blog/2015/07/30/cifar.html}. The shortlist of test images used for evaluation can be downloaded here \url{http://www.cs.toronto.edu/~emansim/cap2im/test-nns.pkl}.} \citep{krizhevsky_cifar}. 
Since there are ``easy'' images for which the model assigns high log-probabilities independent of the query caption, 
we instead look at the ratio of 
the likelihood of the image conditioned on the sentence to the likelihood of the image conditioned on the mean sentence representation in the training set, following the retrieval protocol of~\citep{kiros_captions}.
We also found that the lower bound on the test log-probabilities decreased for sharpened images, and that sharpening considerably hurt the retrieval results. Since sharpening changes the statistics of images, the estimated log-probailities of image pixels is not necessarily a good metric.

>>>>

\section{Experiments}
\vspace{-0.05in}
\subsection{Microsoft COCO}
\vspace{-0.05in}
Microsoft COCO \citep{mscoco} is a very large dataset of roughly 83k images, each annotated with 5 captions. The rich collection of images with a wide variety of styles, backgrounds and objects makes the task of learning a good generative model 
%conditioned on a caption 
very challenging. For consistency with related work on caption generation, we used only the first five captions when training and evaluating our model. 
The images were resized to $32 \times 32$ dimension for consistency with other tiny image datasets \citep{krizhevsky_cifar}. In the following subsections, we analyzed both the qualitative and quantitative aspects of our model as well as compared its performance with that of other, related generative models\footnote{To see more
generated images, visit \url{http://www.cs.toronto.edu/~emansim/cap2im.html}}.

@@

We compared the performance of the proposed model to the DRAW model conditioned on captions without $align$ function (noalignDRAW) as well as the DRAW model conditioned on skipthought vectors of \cite{kiros_skipthought} (skipthoughtDRAW). All of the conditional DRAW models were trained with binary cross-entropy cost function. We also compared our model with Fully-Connected (Fully-Conn) and Convolutional-Deconvolutional (Conv-Deconv) Variational Autoencoders which were trained with the least squares cost function. The LAPGAN model of~\cite{denton_lapgan} was trained on a two level Laplacian Pyramid with a GAN as a top layer generator and all stages were conditioned on the same skipthought vector.

\begin{table}[!t]
\begin{center}
\begin{tabulary}{\linewidth}{c || c c c c c || c c}
\hline
\multicolumn{7}{c}{\textbf{Microsoft COCO (prior to sharpening)}} \\
\hline
& \multicolumn{5}{c||}{Image Retrieval} & \multicolumn{1}{c}{Image Similarity} \\
%\cline{2-6}
\textbf{Model} & \textbf{R@1} & \textbf{R@5} & \textbf{R@10} & \textbf{R@50} & \textbf{Med r} & \textbf{SSI}\\ %& \textbf{Feature Distance} \\
\hline
\hline
LAPGAN & - & - & - & - & - & 0.08 \\ % & 6.89 \\ % - & - & - & - & - & 0.08
\hline
Fully-Conn VAE & 1.0 & 6.6 & 12.0 & 53.4 & 47 & 0.156 \\ %& 6.11 \\ % 0.688 & 4.58 & 8.74 & 46.832 & 54 & -1
Conv-Deconv VAE & 1.0 & 6.5 & 12.0 & 52.9 & 48 & 0.164 \\ %& 6.32 \\ % 0.672 & 4.448 & 8.476 & 46.596 & 54 & 0.131
skipthoughtDRAW & 2.0 & 11.2 & 18.9 & 63.3 & 36 & 0.157 \\ %& 6.93 \\ % 0.988 & 6.18 & 11.48 & 52.512 & 48 & 0.118
noalignDRAW & 2.8 & 14.1 & 23.1 & 68.0 & 31 & 0.155 \\ %& 7.01 \\ % 1.0 & 6.124 & 11.18 & 52.112 & 48 & 0.114
alignDRAW & 3.0 & 14.0 & 22.9 & 68.5 & 31 & 0.156 \\ %& 6.99 \\ % 0.924 & 6.216 & 11.184 & 52.492 & 48 & 0.115
\end{tabulary}
\end{center}
\caption{Retrieval results of different models. 
\textbf{R@K} is Recall@K
    (high is good). \textbf{Med} {\it r} is the median rank (low is
    good). 
}
\label{tab:results}
\vspace{-0.1in}
\end{table}

@@

As for the quantitative comparison of different models, we first compare the performances of the model trained with variational methods. We rank the images in the test set conditioned on the captions based on the variational lower bound of the log-probabilities 
and then report the Precision-Recall metric as an evaluation of the quality of the generative model (see Table~\ref{tab:results}.). Unsurprisingly, generative models did not perform well on the image retrieval task. To deal with the large computational complexity involved in looping through each test image, we create a shortlist of one hundred images including the correct one, based on the images having the closest distance in the convolutional feature-space of a VGG-like model \citep{simonyan_convnet} trained on the CIFAR dataset\footnote{The architecture of the model is described here \url{http://torch.ch/blog/2015/07/30/cifar.html}. The shortlist of test images used for evaluation can be downloaded from \url{http://www.cs.toronto.edu/~emansim/cap2im/test-nns.pkl}.} \citep{krizhevsky_cifar}. 
Since there are ``easy'' images for which the model assigns high log-probabilities independent of the query caption, 
we instead look at the ratio of 
the likelihood of the image conditioned on the sentence to the likelihood of the image conditioned on the mean sentence representation in the training set, following the retrieval protocol of~\citep{kiros_captions}.
We also found that the lower bound on the test log-probabilities decreased for sharpened images, and that sharpening considerably hurt the retrieval results. Since sharpening changes the statistics of images, the estimated log-probailities of image pixels is not necessarily a good metric.
